---
title: "Probability Distribution"
author: "Carlos Amissah"
date: "2024-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

################### true data #############################
```{r}
library(ggplot2) # for graphics
library(MASS)
z <- read.table("fdata.csv",header=TRUE,sep=",")
str(z)
summary(z)

# getting rid of zeros
z <- z[z$tree.density >0,]
str(z)
summary(z$tree.density)


p2 <- ggplot(data=z, aes(x=tree.density, y= ..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2)
print(p2)

# add in a kernel density plot of the data
p2 <-  p2 +  geom_density(linetype="dotted",size=0.75)
print(p2)
```

```{r}
# fit a normal distribution to your data
normPars <- fitdistr(z$tree.density,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```

```{r}
# plot normal probability density
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$tree.density),len=length(z$tree.density))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$tree.density), args = list(mean = meanML, sd = sdML))
p2 + stat
```

```{r}
# exponential probability density
expoPars <- fitdistr(z$tree.density,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$tree.density), args = list(rate=rateML))
p2 + stat + stat2
```

```{r}
# uniform probability density
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$tree.density), args = list(min=min(z$tree.density), max=max(z$tree.density)))
p2 + stat + stat2 + stat3
```

```{r}
# gamma probability density
gammaPars <- fitdistr(z$tree.density,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$tree.density), args = list(shape=shapeML, rate=rateML))
p2 + stat + stat2 + stat3 + stat4
```

```{r}
# beta probability density
pSpecial <- ggplot(data=z, aes(x=tree.density/(max(tree.density + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) +
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$tree.density/max(z$tree.density + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$tree.density), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

```{r}
### simulated data using maximum likely parameters using rgamma as the best fit #############

myGamma <- rgamma(n=454,shape=1.24,rate= 1107.3)
myGamma <- data.frame(1:454, myGamma)
names(myGamma) <- list("ID","myVar")

# histogram plot of data
p1 <- ggplot(data=myGamma, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2)
print(p1)
# adding probability density curve
statp1 <- stat_function(aes(x = myVar, y = ..y..), fun = dgamma, colour="brown", n = length(myGamma$myVar), args = list(shape=1.24, rate=1107.3))
 p1+statp1
```
# comparing the distribution of original data vs simulated data
```{r}
par(mfrow=c(1,2))
p2 + stat4 # original data
p1+statp1 # simulated data
```

How do the two histogram profiles compare? 

The two histogram depicts gamma distribution of density of trees in fields. The first one a real data while the second one is the simulated data. 
Both histograms are right skewed with majority of the data concentrated between 0.00 and 0.002.


Do you think the model is doing a good job of simulating realistic data that match your original measurements? Why or why not?

Yes, the model is did a good job simulating realistic data that matches the original data. The data shows the distribution of tree density in fields. Often, farmers are not motivated to keep trees within fields because trees create a canopy within the field and crops do not grow well under the shade. Farmers remove as many trees within fields to maximize the number seeds sown and also make up space for crops to grow. As a result, you are likely to find one to three trees within a field and the gamma distribution curve captures the picture of tree density in a real world.
