---
title: "Homework_05"
author: "Carlos Amissah"
date: "2024-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```## Question 1
```{r}

n_dims <- sample(x=seq(3,10), size = 1)

n_dims^2

z <- c(1:n_dims^2)

Sample_z <- sample(z)

# create a square matrix
my_m <- matrix(data=Sample_z, ncol=Sample_z, nrow = Sample_z)
print(my_m)

# transpose matrix
t_my_m <- t(my_m)
print(t_my_m)

# calculate sum and mean in 1st row
sum(t_my_m[1,])
mean(t_my_m[1,])
# calculate sum and mean in 1st row
sum(t_my_m[5,])
mean(t_my_m[5,])

eigen(my_m)
# checking kind of numbers
typeof(eigen(my_m)$values)
typeof(eigen(my_m)$vectors)
```
### Question 2
```{r}
# creating a 4*4 matrix
z <- runif(16)
my_matrix <- matrix(data=z, ncol=4)
# creating logical values 
x <- runif(100)
my_logical <- x < 0.47
print(my_logical)
my_letters <- sample(letters)

newlist <- list(my_matrix[2,2], my_logical[2],my_letters[2])

typeof(newlist[1])
typeof(newlist[2])
typeof(newlist[3])

z <- c(0.7743394, TRUE,"S")
typeof(z)
```
### Question 3
```{r}
my_unis<- seq(from=0, to=10, length=26)
my_letters <- sample(LETTERS)

#creating a dataframe 
dframe<- data.frame (my_unis,my_letters)
print(dframe)
# adding NAs randomly to my_unis column 
dframe[sample(length(dframe$my_unis), size = 4),1]<- NA
print(dframe)
# rearranging my_letters in alphabetical order
dframe[order(dframe$my_letters),]
```
